---
title: "week 6 lab"
author: "J.F.P. (Richard) Scholtens s2956586"
date: "Generation date: `r format(Sys.time(), '%b %d, %Y - %H:%M:%S')`"
output: 
  html_document:
    toc: true
    code_folding: show
    toc_float: 
        collapsed: false
        smooth_scroll: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

packages <- c("jtools", "lme4", "visreg")
if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packages, rownames(installed.packages())), repos='http://cran.us.r-project.org') 
}

library(jtools)
library(lme4)
library(visreg)
```

# Load data

```{r load}

mydf = read.csv("lab6_temperature.csv")
str(mydf)


```

# Make Coastal a factor

```{r, fig.width=5, fig.height=5}
str(mydf$Coastal <- as.factor(mydf$Coastal))
```

# Centre variables

```{r, fig.width=5, fig.height=5}
mydf$center.Latitude <- mydf$Latitude - mean(mydf$Latitude)

mydf$center.Longitude <- mydf$Longitude - mean(mydf$Longitude)

mydf$center.Elevation <- mydf$Elevation - mean(mydf$Elevation)

mydf$scale.Elevation <- mydf$center.Elevation/1000

```

# Boxplots

```{r, fig.width=10, fig.height=5}
boxplot(mydf$Temperature ~ mydf$City)
boxplot(mydf$Temperature ~ mydf$Year)
```

We can see there is a lot of variation between the cities. However the variation within the years is almost non existent.

# LMER

```{r, fig.width=5, fig.height=5}
library(lme4)
m.lmer3 <- lmer(Temperature ~ Coastal + center.Latitude + center.Longitude + scale.Elevation + (1|City) + (1|Year), data=mydf, REML=F)
summary(m.lmer3)

```

All the t values are above 2. The minus numbers get an absolute value wich is higher than 2. Therefor all the predictors are significant.

This means the following:

Whenever a city is a coastal city the temperature is about 6.4729 higher.

For every unit of latitude the temperature will drop with -1.8955 degrees.

For every unit of latitude the temperature will increase with 0.2822 degrees.

For every unit of elevation the temperature will drop with -2.1882 degrees.


# Add interactions

```{r, fig.width=5, fig.height=5}
m.lmer4 <- lmer(Temperature ~ Coastal*center.Longitude + Coastal*center.Latitude + Coastal*scale.Elevation + center.Latitude*center.Elevation + center.Longitude*center.Elevation + (1|City) + (1|Year), data=mydf, REML=F)
summary(m.lmer4)$coef
AIC(m.lmer4) - AIC(m.lmer3)
anova(m.lmer3, m.lmer4)

```
We can see a minus integer when we check the AIC. This means the model with the interactions is better. Also the p-level of the anova test is below significance level of 0.05.

There is one interaction which is not significant. The interaction of latitude and elevation has an absolute t-value which is below 2.


# Visualise interactions
```{r, fig.width=10, fig.height=5}
library(visreg)
visreg(m.lmer4, xvar = "center.Latitude", by = "Coastal")
visreg(m.lmer4, xvar = "center.Elevation", by = "Coastal")
```

In the first graph you can see that the temperature drops with more units of latitude for all cities.

I the second graph you can see that the temperature drops quicker for cities which have a high elevation and are a coastal city. However a lot of cities do not have a high elevation. All but 2. This can be seen in the second frame of the graph where you see left a lot of stripes and to the right only two.


# Assumptions (normality)

```{r, fig.width=5, fig.height=5}
qqnorm(resid(m.lmer4))
qqline(resid(m.lmer4))
shapiro.test(resid(m.lmer4))
```

We see a slightly s-shaped line in the plot. This means it in not normally distributed. The shapiro test confirms with a p-level below
the 0.05 signficance lecvel. p-level = 2.2e-16.


# Assumptions (heteroscedasticity)

```{r, fig.width=5, fig.height=5}
plot(fitted(m.lmer4), resid(m.lmer4))
``` 

If look at the dots we can see that all distances on the x-axis look the same. Therefor we can state there is no heteroscedasticity.

# Outliers

```{r, fig.width=5, fig.height=5}
(noutliers <- sum(abs(scale(resid(m.lmer4))) >= 2))
noutliers / nrow(mydf)
```

There are 4 datapoints.
Of these points there are 0.002380952 % of the observations deleted.
