---
title: "lab week5: logistic regression"
author: "J.F.P. (Richard) Scholtens s2956586"
date: "Generation date: `r format(Sys.time(), '%b %d, %Y - %H:%M:%S')`"
output: 
  html_document:
    toc: true
    code_folding: show
    toc_float: 
        collapsed: false
        smooth_scroll: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

packages <- c("rms")
if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packages, rownames(installed.packages())), repos='http://cran.us.r-project.org') 
}

library(rms) #needed for lrm()
```

# Load data

We create a data frame from the table shown in this week's slides (slide 7, do not include column "both").

To avoid typing all the observations (there are 184!) we use the function "c()" to concatenate information and function "rep()" to repeat one element multiple times.

```{r 1load, message=FALSE}
library(rms) #needed for lrm()

# create the variables
status <- c(rep("upper", 36), rep("middle",94), rep("lower",54))
outcome <- c(rep("r", 30), rep("schwa", 6), rep("r", 20), rep("schwa", 74), rep("r", 4), rep("schwa", 50))

# create the data frame
ny.data <- data.frame(outcome, status)
table(ny.data) 
```


# Factor variables

```{r, fig.width=5, fig.height=5}
str(ny.data)
```

As you can see in the results both variables are Factors.

# Reference level

Use the following levels as the reference ones:
- "schwa" in the variable pronunciation
- "lower" in the variable status

```{r, fig.width=5, fig.height=5}
ny.data$outcome <- relevel(ny.data$outcome, ref = "schwa")
levels(ny.data$outcome)
levels(ny.data$status)
```

One of the variables (status) has already a "lower" reference level.


# Fit a model


```{r, fig.width=5, fig.height=5}
(m.lrm <- lrm(outcome ~ status, data = ny.data))
summary(m.glm <- glm(outcome ~ status, data = ny.data, family = binomial))

```

# Assumptions

```{r, fig.width=5, fig.height=5}
rms::vif(m.lrm)
```

We don't have to check the linearity. This is because there are no quantitative exporaty variables.

By using the VIF method we can see there is no multicollinearity because both scores are below the 5.


# Odds of "r" to "schwa" when the predictor status is at its reference level ("lower")

```{r, fig.width=5, fig.height=5}
exp(-2.5257)
```

The intercept was -2.5257. The odds are 0.08000229.


# Probability of the outcome being "r" in this case, i.e. status "lower"?


```{r, fig.width=5, fig.height=5}
exp(-2.5257)/(1+exp(-2.5257))

```

# In which status is one more prone to pronounce "r" (instead of "schwa")? How many times higher are the odds compared to the reference level of status ("lower")?

```{r, fig.width=5, fig.height=5}
exp(4.1352)

```


As we could see in the table in the beginning of this report we see that from status upper has the highest frequency. 


Odds: 62.50209

This means the odds of hearing r over schwa are 62.50209 times more likely in the upper states than in the lower status.

# Confidence intervals

```{r, fig.width=5, fig.height=5}

exp(confint(m.glm))

```

There is no 1 in the intervals. This means they are significant.

# Performance

In the logistic regression model we can see the C level is 0.805. This means it is between acceptable and excellent.
